---
title: "Find the closest pair of points in 2 dimension"
author: "Xiaochuan Yang"
date: "25/09/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Problem

Given a family of $n$ points in 2d square, find the pair that are closest in Euclidean distance and output their distance to each other.

The goal is to complete the task in $O(n\log n)$ time.

# Method

Use the divide and conquer algorithm.

1.  divide the points into left group and right group
2.  find the closest pair in each of the two groups.
3.  find the split pair that are the closest
4.  compare the three distances and output the pair that achieves the minimum

We need to write two functions. The first, "splitPair", finds the closest pair between *two* groups of points, one point from each group. The second, "closestPair", finds the closest pair in *one* group of points.

A naive implementation of this algorithm will take at least $n^2$ time to complete the task, because the split pair function calculate the distance between two groups of points of size $n/2$ each.

Here is how we can avoid this.

1.  divide the points into left group and right group
2.  find the closest pair in each of the two groups, **and also the minimal distance** $\delta$
3.  **find the closest pair among all split pairs within a distance of** $\delta$ to the boundary of the two groups.
4.  compare the three distances and output the pair that achieves the minimum.

Sure, we have abandoned all the points that are not within $\delta$ to the boundary of two groups, thus reduce the running time of step 3, but what if there are still a positive proportion of the $n$ points lying in the $\delta$ neighborhood of the boundary? It turns out that the definition of $\delta$ helps, and a geometric consideration leads to a linear time execution of step 3. This, together with the common sense that a binary tree is of height $\log n$, gives the desired $n\log n$ bounds for the running time of the algorithm.

Now let's do this for real.

# Implementation

let's start.

```{r}

```
